---
title: "Anomaly transformation"
output: html_document
---

```{r setup, include=FALSE}
dir_data <- "D:/HOME/ekamir/scoring_medscope/data/stics" 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure" 

library(easyVerification)
library(verification)
library(dplyr)
library(plyr)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)

no_core <- 6
nboot <- 100
quinzn <- 9:20 # selected quinzaines de [9;20]

```

## Indicator without t dimension

We load values previously extracted and compute ensemble means (mean between runs = models), which is required to compute deterministic scores. Values are stored in 2 3 dimensionial arrays (grids+years+t), one for prevision and one for climatology.
N.B.: 831.22 sec elapsed

```{r, echo=FALSE}

indicateur <- "grassyieldC1C2_"

  cl <- makeCluster(no_core)
  registerDoParallel(cl)
  tictoc::tic()
  foreach(b=1:nboot, .packages="tidyverse") %dopar%  { # loop on bootstrap samples
  
    nom_ref <- paste0(indicateur,b,"_ref.RData")
    nom_prev <- paste0(indicateur,b,"_prev.RData")

    load(file = file.path(dir_data, "extract", nom_ref))
    load(file = file.path(dir_data, "extract", nom_prev))

    ngrids <- dim(array_obs)[1]

    # CROSS VALIDATED ANOMALIES: we compute and save arrays of anomalies using cross validation scheme
    
    prev_anomaly <- array(NA,dim=c(ngrids,nan,nrun))
    obs_anomaly <- array(NA,dim=c(ngrids,nan))
    
    for (y in 1:nan){
      cv_mean_prev <- apply(array_prev[,-y,], c(1,3), mean)
      cv_mean_obs <- apply(array_obs[,-y], 1, mean)
  
      prev_anomaly[,y,] <- array_prev[,y,]-cv_mean_prev
      obs_anomaly[,y] <- array_obs[,y]-cv_mean_obs
    
  }

  nom_prev <- paste0(indicateur,b,"_anomaly_prev.RData")
  nom_obs <- paste0(indicateur,b,"_anomaly_obs.RData")

  save(prev_anomaly, file = file.path(dir_data, "extract", nom_prev))
  save(obs_anomaly, file = file.path(dir_data, "extract", nom_obs))
  
}
stopCluster(cl)

tictoc::toc() 

```
## Indicator with t dimension

We load values previously extracted and compute ensemble means (mean between runs = models), which is required to compute deterministic scores. Values are stored in 2 3 dimensionial arrays (grids+years+t), one for prevision and one for climatology.
N.B.: 831.22 sec elapsed

```{r, echo=FALSE}

indicateur <- "bilanhydr"

  cl <- makeCluster(no_core)
  registerDoParallel(cl)
  tictoc::tic()
  foreach(b=1:nboot, .packages="tidyverse") %dopar%  { # loop on bootstrap samples
  
    nom_ref <- paste0(indicateur,b,"_ref.RData")
    nom_prev <- paste0(indicateur,b,"_prev.RData")

    load(file = file.path(dir_data, "extract", nom_ref))
    load(file = file.path(dir_data, "extract", nom_prev))

    ngrids <- dim(array_obs)[1]
    ndimT <- dim(array_obs)[3]

    # CROSS VALIDATED ANOMALIES: we compute and save arrays of anomalies using cross validation scheme
    
    prev_anomaly <- array(NA,dim=c(ngrids,nan,nrun))
    obs_anomaly <- array(NA,dim=c(ngrids,nan))
    
    prev_seas <- apply(array_prev[,,,quinzn], c(1,2,3), sum) #### aggregation method to be changed!! - TIME PERIOD SELECTION!!!!
    ref_seas <- apply(array_obs[,,quinzn], c(1,2), sum)

    for (y in 1:nan){
      cv_mean_prev <- apply(prev_seas[,-y,], c(1,3), mean)
      cv_mean_obs <- apply(ref_seas[,-y], 1, mean)
  
      prev_anomaly[,y,] <- prev_seas[,y,]-cv_mean_prev
      obs_anomaly[,y] <- ref_seas[,y]-cv_mean_obs
    
  }

  nom_prev <- paste0(indicateur,b,"_anomaly_prev.RData")
  nom_obs <- paste0(indicateur,b,"_anomaly_obs.RData")

  save(prev_anomaly, file = file.path(dir_data, "extract", nom_prev))
  save(obs_anomaly, file = file.path(dir_data, "extract", nom_obs))
  
}
stopCluster(cl)

tictoc::toc() 

```
