---
title: "Exploration"
output: html_document
---
The code aims at creating 2 types of plot for data visualisation
It requires data extracted by "extraction"
########## ADD LEGENDS !!!!!!!!!!!!!!!!!! ########################

```{r setup, include=FALSE}

library(s2dverification)
library(stats)
library(s2dverification)
library(ggplot2)
library(CSTools)
library(fanplot)
library(dplyr)
library(ggfan)

```

## Parameters 

Some are loaded (because they remain permanent through codes)
Some are set here and can therefore be modified by user

```{r }

                                     ### SELECTED PARAMETERS (to be set by user) ###


# selecting project
selected_project <- "stics" # stics or season or tree
# directories and cores
setup(project = selected_project, no_cores = 6)

# selecting variable of interest
load(file = file.path(dir_data, "extract", "indicateurs.RData"))
indicateur <- indicateurs[4] 

# selecting grid of interest (for second plot)
maille <- 10


                                      ### METADATA ###


# loading number of bootstrap interations
load(file = file.path(dir_data, "extract", "nboot.RData"))

# loading time scales and agregation type
load(file = file.path(dir_data, "extract", "temporalite.RData")) # is there a time dimension
load(file = file.path(dir_data, "extract", "time_window.RData")) # if yes which window is selected
load(file = file.path(dir_data, "extract", "agregation.RData")) # if yes what type of aggregation should be applied
ntemp <- length(temp) # number of elements in time dimension

# loading ensemble members
load(file = file.path(dir_data, "extract", "runs.RData"))
nrun <- length(run) # number of members

# loading years
load(file = file.path(dir_data, "extract", "years.RData"))
nan <- length(an) # number of years

 
```

## Retrieving original data set

To explore original data we need to reshape it, because bootstrap has modified year order (random sampling doesn't respect chronology).

```{r }

  d <- which(indicateurs==indicateur)
  temporalite <- time_scale[d] # select corresponding time dimension: single or multi
  agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean
  
  # getting number of grids automatically
  load(file = file.path(dir_data, "extract", paste0(indicateur, "_lon.RData")))
  ngrids <- length(lon)
  

                                            ### WITH TIME DIMENSION ###
  
if (temporalite=="multi"){
  
# creating arrays for original data to be stored in
ps_original <- array(NA, dim=c(ngrids, nan, nrun, ntemp))
ref_original <- array(NA, dim=c(ngrids, nan, ntemp))

 for (year in an){
 y <- which(an==year)

    # loading list of bootstrapped years 
    load(file = file.path(dir_data, "extract", "rand_years.RData")) 
     
    list_match <- lapply(rand_year, function(x) which(year == x)) # list of matchs between list of years randomly generated for the bootstrap and the year we are interested in
     
     y_index <- list_match[[1]] # year index within the bootstrap iteration
     b_index <- 1 # bootstrap iteration index
     for (b in 1:(nboot-1)){ # pick the first iteration and first index matching the year we are looking for
       if (length(list_match[[b]])<1 & length(y_index)<1) {y_index <- list_match[[b+1]]; b_index <- b+1}
       else {y_index <- y_index; b_index <- b_index}
     }
     
  # loading ps bootstrapped years
  nom_prev <- paste0(indicateur,b_index,"_prev_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_prev))
  # we reshape data following original chronology
  ps_original[,y,,] <- array_prev[,y_index[1],,temp]
  
  # loading ps bootstrapped years
  nom_ref <- paste0(indicateur,b_index,"_ref_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_ref))
  # we reshape data following original chronology
  ref_original[,y,] <- array_obs[,y_index[1],temp]
}


                                            ### WITHOUT TIME DIMENSION ###
# same as before, without the time dimension

} else if (temporalite=="single"){

# creating arrays for original data to be stored in
ps_original <- array(NA, dim=c(ngrids, nan, nrun))
ref_original <- array(NA, dim=c(ngrids, nan))

 for (year in an){
 y <- which(an==year)

       # bootstrap iteration index of the year selected
    load(file = file.path(dir_data, "extract", "years.RData")) # list of years in each bootstrap iteration
     
    list_match <- lapply(rand_year, function(x) which(year == x)) # list of matchs between list of years randomly generated for the bootstrap and the year we are interested in
     
     y_index <- list_match[[1]] # year index within the bootstrap iteration
     b_index <- 1 # bootstrap iteration index
     for (b in 1:(nboot-1)){ # pick the first iteration and first index matching the year we are looking for
       if (length(list_match[[b]])<1 & length(y_index)<1) {y_index <- list_match[[b+1]]; b_index <- b+1}
       else {y_index <- y_index; b_index <- b_index}
     }
     
  nom_prev <- paste0(indicateur,b_index,"_prev_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_prev))
  ps_original[,y,] <- array_prev[,y_index[1],]
  
  nom_ref <- paste0(indicateur,b_index,"_ref_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_ref))
  ref_original[,y] <- array_obs[,y_index[1]]
}

}
```

## Data aggregation

We compute temporal and ensemble aggregation needed for the plots
- grid means are needed for boxplot
- ensemble means are needed for the probabilistic plot

```{r }


                                            ### WITH TIME DIMENSION ###
if (temporalite=="multi"){
      if (agreg=="sum"){
      ps_agreg <- apply(ps_original[,,,], c(1,2,3), sum)
      ref_agreg <- apply(ref_original[,,], c(1,2), sum)
      } else if (agreg=="mean"){
      ps_agreg <- apply(ps_original[,,,], c(1,2,3), mean)
      ref_agreg <- apply(ref_original[,,], c(1,2), mean)
      }
      
ps_gridmean <- apply(ps_agreg, c(2,3), mean)
ps_ensmean <- apply(ps_agreg, c(1,2), mean)
ps_original <- ps_agreg
ref_gridmean <- apply(ref_agreg, 2, mean)
ref_original <- ref_agreg



                                            ### WITHOUT TIME DIMENSION ###
} else if (temporalite=="single"){

ps_gridmean <- apply(ps_original, c(2,3), mean)
ps_ensmean <- apply(ps_original, c(1,2), mean)
ref_gridmean <- apply(ref_original, 2, mean)
}

```

## BOXPLOT AND PROB SERIES

Attention sum ou mean selon indicateur!

```{r }

# re shaping data: we put them in a table with 3 columns: type(ref or prevision) - year - value 
df_raw <- data.frame() 

for (y in 1:nan){

  df_raw <- rbind(df_raw, data.frame(type = "prevision", year=an[y], value = ps_gridmean[y,]))
  df_raw <- rbind(df_raw, data.frame(type = "ref", year=an[y], value = ref_gridmean[y]))

}

df_raw$year <- as.factor(df_raw$year)


# plotting with ggplot2

p1<-ggplot() + geom_boxplot(df_raw %>% filter(type=="prevision"), mapping=aes(x=year, y=value), color="#dd1c77", show.legend = TRUE)+
  geom_point(df_raw %>% filter(type=="ref"), mapping=aes(x=year, y=value), color="#2ca25f")+
  #scale_colour_manual(name="Error Bars",values=cols) + scale_fill_manual(name="Bar",values=cols)
  labs(title = paste0(indicateur, scale, " spatial means"), x="Year", y= "Value")+
      theme_bw()+
       scale_x_discrete(breaks = scales::pretty_breaks(n = 12))

#```

## Plot ensemble mean and spread vs observation

#```{r }


df_psmean <- data.frame() # table with: type - year - value(ensemble mean)
df_quant <- data.frame() # table with: type - year - value(ensemble quantiles)
df_ref <- data.frame() # table with: type - year - value(ensemble quantiles)

for (y in 1:nan){
  
  df_quant <- rbind(df_quant, data.frame(year=an[y], value = ps_original[maille, y, ]))
  df_psmean <- rbind(df_psmean, data.frame(year=an[y], value = ps_ensmean[maille, y]))
  df_ref <- rbind(df_ref, data.frame(year=an[y], value = ref_original[maille, y]))
}

# reference terciles
ter1_ref <- as.numeric(quantile(ref_original[maille, ], probs = 1/3))
ter3_ref <- as.numeric(quantile(ref_original[maille, ], probs = 2/3))

# ps terciles
ter1_ps <- as.numeric(quantile(ps_original[maille,,], probs = 1/3))
ter3_ps <- as.numeric(quantile(ps_original[maille,,], probs = 2/3))

# plot with ggplot2
p <- ggplot(df_quant, aes(x=year, y=value)) + 
  labs(title = paste0(indicateur, " MJJASO grid ", maille), x="Year", y= "Value")+
      theme_bw()+
       scale_x_continuous(breaks = scales::pretty_breaks(n = 12))


## with geom_fan extra
p2<- p+geom_fan(data=df_quant, intervals=c(1,0.1), show.legend = FALSE) + 
  geom_line(data=df_ref, mapping=aes(x=year,y=value), color="#a1d99b", size=1, show.legend = TRUE) +
  geom_hline(yintercept=ter1_ref, linetype="dashed", color = "#a1d99b", size=0.4) +
  geom_hline(yintercept=ter3_ref, linetype="dashed", color = "#a1d99b", size=0.4) +
  geom_line(data=df_psmean, mapping=aes(x=year,y=value), color="#f03b20", size=1) +
  geom_hline(yintercept=ter1_ps, linetype="dashed", color = "#f03b20", size=0.4) +
  geom_hline(yintercept=ter3_ps, linetype="dashed", color = "#f03b20", size=0.4)

pdf(paste0(dir_plot,"/explo_",indicateur,scalename(selected_project,temporalite,agreg),"ps.pdf"))
par(mfrow=c(1,2))
print(p1)
print(p2)
dev.off()



```
