---
title: "data_extraction"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
dir_data <- "D:/HOME/ekamir/scoring_medscope/data/stics" 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure" 

library(plyr)
library(dplyr)
library(tidyverse)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)

```

## Parameters 

```{r }

# bootstrap and cores
no_core <- 6
nboot <- 100

# variables
indicateurs <- c("drainall", "etall", "yieldwheat", "yieldgrass", "yieldgrassC2C3")
  ### temporalité
  if (indicateur=="yieldwheat"|indicateur=="yieldgrass"|indicateur=="yieldgrassC2C3"){
      temporalite <- "yearly"
  }else {temporalite <- "quinzaine"}

# runs
run <- seq(from=1, to=25, by=1)
nrun <- length(run)

# temporal agregation
temp <- 9:20
ntemp <- length(temp)

```

## Loading data in csv file

```{r data loading, echo=FALSE}

# UERRA

raw_drainall <- read.csv(file = file.path(dir_data, "Extract_STICS-KCETP_DRAIN_ALL_P05_crau45_1993_2016.csv"), header = TRUE, sep = ",")

raw_etall <- read.csv(file = file.path(dir_data, "Extract_STICS-KCETP_ET_ALL_P05_crau45_1993_2016.csv"), header = TRUE, sep = ",")

raw_yieldwheat <- read.csv(file = file.path(dir_data, "Extract_STICS-KCETP_RENDEMENT_WHEAT_P05_crau45_1993_2016.csv"), header = TRUE, sep = ",")

raw_yieldgrass <- read.csv(file = file.path(dir_data, "Extract_STICS-KCETP_RENDEMENT_GRASS_P05_crau45_1993_2016.csv"), header = TRUE, sep = ",")

raws <- list()
raws[[1]] <- raw_drainall
raws[[2]] <- raw_etall
raws[[3]] <- raw_yieldwheat
raws[[4]] <- raw_yieldgrass %>% filter(period == "ANNEE") 
raws[[5]] <- raw_yieldgrass %>% filter(period == "C2" | period == "C3") %>% group_by(climattype, climatscenario, year_, cellnum) %>% mutate(indicatorvalue=sum(indicatorvalue)) %>% filter(period == "C2")

```

## Automatic parameters extraction

List of variables, years, grids, extracted directly from the data set.

```{r indicators, echo=FALSE}

# years
an <- unique(raw_drainall[,"year_"])
nan <- length(an)

# temporal dimension
dimT <- unique(raw_drainall[,"period"])
ndimT <- length(dimT)

```

## Extracting indicator values

For each indicator, values are stored in 3 multi-dimensional arrays: one per climat type (reference, prevision, climatology)
Reference array has 3 dimension: grid+year+time
Prevision/climatology arrays have 4 dimensions: grid+year+run+time

```{r, echo=FALSE}

# generating random year series
# rand_year <- list()
# for (b in 1:nboot){
#    rand_year[[b]] <- sample(1:nan, nan, replace = TRUE) 
#  }
#  save(rand_year, file = file.path(dir_data, "extract", "years.RData"))

load(file = file.path(dir_data, "extract", "years.RData")) # list of years in each bootstrap iteration

tictoc::tic() 

for (d in 1:3) {
  
  data <- raws[[d]]
  indicateur <- indicateurs[d]
  
                                               ### COORDINATES ###
  
  
  # we pick a random climat type, year and t dimension, as grid range assessed remains the same no matter what are the values of these parameters

  # longitudes
raw_filtered <- raws[[d]] %>% select(cellnum,lon,lat) %>% distinct(cellnum, .keep_all= TRUE)
lon <- raw_filtered %>% select(lon)
lon <- as.numeric(unlist(lon))

save(lon, file = file.path(dir_data, "extract", paste0(indicateur, "_lon.RData")))

  # latitudes
lat <- raw_filtered %>% select(lat)
lat <- as.numeric(unlist(lat))

save(lat, file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))

  # number of grids
ngrids <- dim(raw_filtered)[1]

 # climat type filtering
    obs_data <- data %>% 
      filter(climattype == "REF") 
    
    prev_data <- data %>% 
      filter(climattype == "PS") 
    
    clim_data <- data %>% 
      filter(climattype == "CLIMATO") 

 # extracting data
cl <- makeCluster(no_core)
registerDoParallel(cl)

foreach(b=1:nboot, .packages="tidyverse") %dopar%  { # loop on bootstrap samples
  
      
                                      ### EXTRACTING VARIABLES WITHOUT TIME DIMENSION ###
      
  if (temporalite=="yearly"){
    
    array_obs <- array(NA, dim=c(ngrids, nan)) 
    array_prev <- array(NA, dim=c(ngrids, nan, nrun))
    array_clim <- array(NA, dim=c(ngrids, nan, nrun))

    for (j in 1:nan){
      y <- rand_year[[b]][j] # we pick a year from the randomize serie
      
      # at first we filter on year
      obs_int <- obs_data %>% 
      filter(year_ == y) 
      prev_int1 <- prev_data %>% 
      filter(year_ == y) 
      clim_int1 <- clim_data %>% 
      filter(year_ == y) 


        array_obs[,j] <- as.vector(unlist(obs_int %>% arrange(cellnum) %>% select(indicatorvalue))) 
 
      # finally filtering on runs (prevision and climatology only)
        for (r in 1:nrun){
        array_prev[,j,r] <- as.vector(unlist(prev_int1 %>% filter(climatscenario == r) %>% arrange(cellnum) %>% select(indicatorvalue)))
        array_clim[,j,r] <- as.vector(unlist(clim_int1 %>% filter(climatscenario == r) %>% arrange(cellnum) %>% select(indicatorvalue)))
        }

    }
    
  
                                      ### EXTRACTING VARIABLES WITH TIME DIMENSION ###
    
       # creating arrays to store data: grids*years(*run)*time 
        
  } else if (temporalite=="quinzaine") { array_obs <- array(NA, dim=c(ngrids, nan, ndimT)) 
    array_prev <- array(NA, dim=c(ngrids, nan, nrun, ndimT))
    array_clim <- array(NA, dim=c(ngrids, nan, nrun, ndimT))

    for (j in 1:nan){
      y <- rand_year[[b]][j] # we pick a year from the randomize serie
      
      # at first we filter on year
      obs_int <- obs_data %>% 
      filter(year_ == y) 
      prev_int1 <- prev_data %>% 
      filter(year_ == y) 
      clim_int1 <- clim_data %>% 
      filter(year_ == y) 


      # then we filter on time dimension (within a year)
      for (t in c(1:ndimT)){  
        array_obs[,j,t] <- as.vector(unlist(obs_int %>% filter(period == dimT[t]) %>% arrange(cellnum) %>% select(indicatorvalue))) 
        
        prev_int2 <- prev_int1 %>% 
        filter(period == dimT[t])  
        clim_int2 <- clim_int1 %>% 
        filter(period == dimT[t])  
 
      # finally filtering on runs (prevision and climatology only)
        for (r in 1:nrun){
        array_prev[,j,r,t] <- as.vector(unlist(prev_int2 %>% filter(climatscenario == r) %>% arrange(cellnum) %>% select(indicatorvalue)))
        array_clim[,j,r,t] <- as.vector(unlist(clim_int2 %>% filter(climatscenario == r) %>% arrange(cellnum) %>% select(indicatorvalue)))
        }
      }
    }
      
    }
  
    nom_ref <- paste0(indicateur,b,"_ref_uerra5.RData")
    nom_prev <- paste0(indicateur,b,"_prev_uerra5.RData")
    nom_clim <- paste0(indicateur,b,"_clim_uerra5.RData")
    
    # saving data
    save(array_obs, file = file.path(dir_data, "extract", nom_ref))
    save(array_prev, file = file.path(dir_data, "extract", nom_prev))
    save(array_clim, file = file.path(dir_data, "extract", nom_clim))
  }
  
} 

stopCluster(cl)


tictoc::toc()

```
