---
title: "map_brier"
output: html_document
---
The code creates a map of AUC score on a selected variable and quantile. It requires scores computed in "score_auc". Signficance of the Brier skill score is represented with dots (it compares ps to climato).

```{r setup, include=FALSE}

library(maps)
library(nnet)
library(ggplot2)
library(latticeExtra)
library(sp)
library(rgdal)
library(sf)
library(ggmap)
library(maps)
library(ggplot2)
library(latticeExtra)
library(sp)
library(rgdal)
library(sf)
library(ggmap)
library(mapview)
library(ggnewscale)

```

## Parameters 

Some are loaded (because they remain permanent through codes)
Some are set here and can therefore be modified by user

```{r}

                                     ### SELECTED PARAMETERS (to be set by user) ###

# selecting project
selected_project <- "season" # stics or season or tree
# directories and cores
setup(project = selected_project, no_cores = 6)

# selecting variable of interest
load(file = file.path(dir_data, "extract", "indicateurs.RData"))
nindicateur=5
indicateur <- indicateurs[nindicateur]

# chosen quantile
quant <- 4 # 4 = low , 6 = upper 
extremes <- "first" # "first" if low; "last" otherwise
kantile <- "ter" # ter or quint

# score value interval: to be modified, especially if all grids are black, it means they're out of range
bss_limits <- c(0,1)

# error limit
na_limit <- 10

# color palette for acc score
score_palette <- c("#ca0020", "#f4a582", "#f7f7f7", "#92c5de", "#0571b0")

# !! DO NOT MODIFY !! one color palette for signifiance plotting  
sign_palette <- c("black", "black")

                                      ### METADATA ###

# loading time scales and agregation type
load(file = file.path(dir_data, "extract", "temporalite.RData")) # is there a time dimension
load(file = file.path(dir_data, "extract", "time_window.RData")) # if yes which window is selected
load(file = file.path(dir_data, "extract", "agregation.RData")) # if yes what type of aggregation should be applied

# trick to get automatically number of grids 
nom_ref <- paste0(indicateur,1,"_ref_uerra5.RData")
load(file = file.path(dir_data, "extract", nom_ref))
ngrids <- dim(array_obs)[1]

# grid size (depending on the zoom)
if (selected_project=="season"|selected_project=="tree"){
  grid_size <- 0.85
} else if (selected_project=="stics"){
  grid_size <- 2.15
}

# background map
if (selected_project=="season"|selected_project=="tree"){
    back_map <- get_map(location = c(1.2, 41, 10, 46), maptype = "satellite", source = "osm") 
} else if (selected_project=="stics"){
    back_map <- get_map(location = c(4, 43.1, 5.7, 44), maptype = "satellite", source = "osm") 
}

```

## Mapping Brier score through a shapefile

We transform the array of scores into a shapefile containing gridded score values and their significance.
Significance is computed using Wilcoxon's test (H0: mean (bss ps - bss clim) value not above 0) on bootstrapped score values.

```{r }

for (indicateur in indicateurs){

                # BUIDLING SHAPEFILE OF MEAN AUC VALUES AND THEIR SIGNIFICANCE #

d <- which(indicateurs==indicateur)
temporalite <- time_scale[d] # select corresponding time dimension: single or multi
agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean

  ## extracting bootstrapped brier score values
# setting up arrays for all bootstrapped scores to be stored in
score_boot <- array(NA, dim=c(ngrids, 10, nboot)) 

# loading data and filling up the array just set
  for (b in 1:nboot)  { 
  
    nom_scoreroc <- paste0(indicateur,"_", b, scalename(selected_project,temporalite,agreg), kantile, "aucgrid.RData")
    load(file = file.path(dir_scores, nom_scoreroc))

    score_boot[,,b] <- scoreprob[,]

  }

  ## we build a matrix gathering all info needed for the shapefile:

    # coordinates

load(file = file.path(dir_data, "extract", paste0(indicateur, "_lon.RData")))
load(file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))
  
matrix4shp <- matrix(0, nrow = ngrids, ncol = 4) # col: skill score, significant, lat, lon 
matrix4shp[,3] <- lat
matrix4shp[,4] <- lon
  
    # mean skill and wilcoxon test outcome

matrix4shp[,2] <- rep(0, ngrids) # signficance: 0 everywhere first, will switch to 1 if positive statistical test outcome
matrix4shp[,1] <- apply(score_boot[,quant+index_clima,], 1, mean, na.rm = TRUE) #  mean of bootstrapped auc  

    # Wilcoxon test

for (i in 1:ngrids)  { 
    if (length(which(is.na(score_boot[i,quant+index_clima,]))) >= 25) {matrix4shp[i,2] <- 0; matrix4shp[i,1] <- NA}
    else { wbrier <- wilcox.test(score_boot[i,quant+index_clima,], mu = 0.3, alternative = "greater", conf.level = 0.95) # h0 represented by "mu = 0" & "alternative = greater", alternative being h1

      if (wbrier$p.value < 0.05) { # p value de 5%
      matrix4shp[i,2] <- 1 # 1 = test validé, 0 = non validé (voir plus haut)
      }
    }
}
    # building shapefile 

df_shp <- data.frame(matrix4shp)
colnames(df_shp) <- c("bss", "significance","lat","lon")
coordinates(df_shp) <- ~lon+lat
proj4string(df_shp) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
shp_grid <- st_as_sf(df_shp) # we transform spatial data frame into shapefile, which is plotting friendly
shp_grid$significance[which(shp_grid$significance==0)] <- NA # to help plotting, we set significance as NA instead of 0


                                      # MAPPING WITH GGPLOT2 #
pdf(paste0(dir_plot,"/brier_",extremes,kantile,"_",indicateur,scalename(selected_project,temporalite,agreg),climat_select,".pdf"))

p<-ggmap(back_map) +
geom_sf(data = shp_grid, aes(color = bss), size=grid_size, shape = 15) + 
  scale_color_gradientn(colours=score_palette, limits = bss_limits, na.value = "black", guide = guide_colorbar(barwidth = 10, barheight = 0.25, nbin = 5)) +
new_scale("color")+
geom_sf(data = shp_grid, aes(color = significance), shape = 3, size = 0.15, show.legend = FALSE) +
  scale_color_gradientn(colours=sign_palette, na.value = "transparent")+
theme(legend.direction = "horizontal", legend.position = "right", legend.title = element_text(), legend.text=element_text(size=6),  
      plot.title = element_text(size = 12, hjust = 0, vjust = 1.5)) + 
  labs(title = paste0(indicateur, scalename(selected_project,temporalite,agreg), extremes, "_", kantile,"_BRIER_", climat_select))
print(p)

dev.off()

}

```
