---
title: "Mapping RMSE"
output: html_document
---
The code creates a map of RMSE score on a selected variable. It requires scores computed in "score_deterministic". Signficance of the score (rmse(ps)-rmse(clim)) is represented with dots.

```{r setup, include=FALSE}

library(maps)
library(nnet)
library(ggplot2)
library(latticeExtra)
library(sp)
library(rgdal)
library(sf)
library(ggmap)
library(maps)
library(ggplot2)
library(latticeExtra)
library(sp)
library(rgdal)
library(sf)
library(ggmap)
library(mapview)
library(ggnewscale)

```

## Parameters 

Some are loaded (because they remain permanent through codes)
Some are set here and can therefore be modified by user

```{r setup, include=FALSE}

                                     ### SELECTED PARAMETERS (to be set by user) ###

# selected project and directories
selected_project <- "season" # stics or season or tree
dir_data <- paste0("D:/HOME/ekamir/scoring_medscope/data/",selected_project) 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure" 

# selecting variable of interest
load(file = file.path(dir_data, "extract", "indicateurs.RData"))
indicateur <- indicateurs[3]

# time integration
scale <- " MJJASO"

# error limit
na_limit <- 10

# color palette for acc score
score_palette <- c("white", "#fee5d9", "#fcbba1", "#fc9272", "#fb6a4a", "#e31a1c")

# !! DO NOT MODIFY !! one color palette for signifiance plotting  
sign_palette <- c("black", "black")

                                      ### LOADED PARAMETERS (permanent) ###


# trick to get automatically number of grids 
nom_ref <- paste0(indicateur,1,"_ref_uerra5.RData")
load(file = file.path(dir_data, "extract", nom_ref))
ngrids <- dim(array_obs)[1]

# grid size (depending on the zoom)
if (selected_project=="season"|selected_project=="tree"){
  grid_size <- 0.85
} else if (selected_project=="stics"){
  grid_size <- 2.15
}

# background map
key = "AIzaSyCD2-pP0yfM8aC8sjCB4g--CaCcExP6NhU"; "AIzaSyB24WUzCzX1lErLgsVLtio1c7QxLfw1_SA"
register_google(key = "AIzaSyB24WUzCzX1lErLgsVLtio1c7QxLfw1_SA")
if (selected_project=="season"|selected_project=="tree"){
    back_map <- get_map(location = c(1.2, 41, 10, 46), maptype = "satellite", source = "google") 
} else if (selected_project=="stics"){
    back_map <- get_map(location = c(4, 43.1, 5.7, 44), maptype = "satellite", source = "google") 
}

```


## Creating map through shapefile and ggplot2 package

We transform arrays of scores into a shapefile containing gridded score values (averaged over bootstrap) and their significance.
Significance is computed using Wilcoxon's test (H0: mean (rmse ps - remse clim) value not above 0) on bootstrapped score values.

```{r }

                # BUIDLING SHAPEFILE OF MEAN RMSE VALUES AND THE SIGNIFICANCE OF RMSE(ps)-RMSE(clim) #

  ## extracting bootstrapped rmse values
# setting up arrays for all bootstrapped scores to be stored in
score_ps <- array(NA, dim=c(ngrids, 3, nboot)) 

# loading data and filling up the array just set
  for (b in 1:nboot)  { 
  
    nom_det <- paste0(indicateur, b,"scoredet_grid_uerra5.RData")
    load(file = file.path(dir_data, "score", nom_det))

    score_ps[,,b] <- score_det[,]

  }

  ## we build a matrix gathering all info needed for the shapefile:

    # coordinates

load(file = file.path(dir_data, "extract", paste0(indicateur, "_lon.RData")))
load(file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))
  
matrix4shp <- matrix(0, nrow = ngrids, ncol = 4) # col: score, significant, lat, lon 
matrix4shp[,3] <- lat
matrix4shp[,4] <- lon
  
    # mean skill and wilcoxon test outcome

matrix4shp[,2] <- rep(0, ngrids) # signficance: 0 everywhere first, will switch to 1 if positive statistical test outcome
matrix4shp[,1] <- apply(score_ps[,2,], 1, mean, na.rm = TRUE) #  mean of bootstrapped RMSE 

    # Wilcoxon test

for (i in 1:ngrids)  { 
  if (length(which(is.na(score_ps[i,2,]))) >= na_limit) {matrix4shp[i,2] <- 0; matrix4shp[i,1] <- NA}
  else {wmsess <- wilcox.test(score_ps[i,3,]-score_ps[i,2,], mu = 0, alternative = "greater", conf.level = 0.95) # h0 represented by "mu = 0" & "alternative = greater", alternative being h1. We assess RMSE ps > RMSE clim

    if (wmsess$p.value < 0.05) { # p value 5%
      matrix4shp[i,2] <- 1 # 1 = test validé, 0 = non validé (voir plus haut)
    }
  }
}

    # building shapefile 

df_shp <- data.frame(matrix4shp)
colnames(df_shp) <- c("rmse", "significance","lat","lon")
coordinates(df_shp) <- ~lon+lat
proj4string(df_shp) <- CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
shp_grid <- st_as_sf(df_shp) # we transform spatial data frame into shapefile, which is plotting friendly
shp_grid$significance[which(shp_grid$significance==0)] <- NA # to help plotting, we set significance as NA instead of 0


                                      # MAPPING WITH GGPLOT2 #

ggmap(back_map) +
geom_sf(data = shp_grid, aes(color = rmse), size=grid_size, shape = 15) + 
  scale_color_gradientn(colours=score_palette,  na.value = "black", guide = guide_colorbar(barwidth = 10, barheight = 0.5, nbin = 5)) + 
new_scale("color")+
geom_sf(data = shp_grid, aes(color = significance), shape = 3, size = 0.15, show.legend = FALSE) +
  scale_color_gradientn(colours=sign_palette, na.value = "transparent")+
theme(legend.direction = "horizontal", legend.position = "right", legend.title = element_text(),  
      plot.title = element_text(size = 12, hjust = 0, vjust = 1.5)) + 
  labs(title = paste0(indicateur,scale," RMSE "))

```
