---
title: "Area under ROC"
output: html_document
---
The code aims at computing probabilistic scores, being Area Under Curve of the first and last quantile, Brier scores + skills scores of the first and last quantile as well.
It is computed for prevision only, except Brier skill score which compare ps and climato: BSS = 1-BS(ps)/BS(clim)
This requires both "raw" data extracted by "extraction" and  probabilistic data computed by "transfo_probin"

```{r setup, include=FALSE}

library(easyVerification)
library(verification)
library(dplyr)
library(stats)
library(SpecsVerification)
library(s2dverification)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)
library(readr)
library(purrr)

```

## Parameters 

Some are loaded (because they remain permanent through codes)
Some are set here and can therefore be modified by user

```{r }

                                     ### SELECTED PARAMETERS (to be set by user) ###

# selecting project
selected_project <- "season" # stics or season or tree
# directories and cores
setup(project = selected_project, no_cores = 6)

# chosen quantile
kantile <- "ter" # ter or quint
if (kantile=="quint"){
  proba = 1:4/5; events = 5
} else { proba = 1:2/3; events = 3}


                                      ### METADATA ###


# loading number of bootstrap interations
load(file = file.path(dir_data, "extract", "nboot.RData"))

# loading variables♣¥
load(file = file.path(dir_data, "extract", "indicateurs.RData"))
# loading time scales and agregation type
load(file = file.path(dir_data, "extract", "temporalite.RData")) # is there a time dimension
load(file = file.path(dir_data, "extract", "time_window.RData")) # if yes which window is selected
load(file = file.path(dir_data, "extract", "agregation.RData")) # if yes what type of aggregation should be applied

ntemp <- length(temp) # number of elements in time dimension

# loading ensemble members
load(file = file.path(dir_data, "extract", "runs.RData"))
nrun <- length(run) # number of members

# loading years
load(file = file.path(dir_data, "extract", "years.RData"))
nan <- length(an) # number of years


```
                                            
## Gridded scores

Scores computed for each grid separately.
They are stored in an array, in the following order(on the 2nd dimension):
1: roc quant1 ps
2: roc quant1 clim
3: roc last quant ps
4: roc last quant clim
5: brier quant1 ps
6: brier quant1 clim
7: brier last quant ps
8: brier last quant clim
9: brier ss quant1 (ps vs clim)
10: brier ss last quant (ps vs clim)

```{r}

tictoc::tic() # start timer
for (indicateur in indicateurs){

  d <- which(indicateurs==indicateur)
  temporalite <- time_scale[d] # select corresponding time dimension: single or multi
  agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean

cl <- makeCluster(no_core) # setting up number of cores
registerDoParallel(cl) # parallelizing on them

foreach(b=1:nboot, .packages = c('easyVerification', 's2dverification'), .combine=rbind) %dopar%  { # for-loop parallelized

  # loading "raw" and probabilistic data
  nom_ref <- paste0(indicateur,b,"_ref_uerra5.RData")
  nom_prev <- paste0(indicateur,b,"_prev_uerra5.RData")
  nom_clim <- paste0(indicateur,b,"_clim_uerra5.RData")
  nom_prev_prob <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"psprob",kantile, ".RData")
  nom_clim_prob <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"climprob",kantile, ".RData")
  nom_ref_bin <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"refbin",kantile, ".RData")
 
  load(file = file.path(dir_data, "extract", nom_prev))
  load(file = file.path(dir_data, "extract", nom_clim))
  load(file = file.path(dir_data, "extract", nom_ref))
  load(file = file.path(dir_data, "extract", nom_prev_prob))
  load(file = file.path(dir_data, "extract", nom_clim_prob))
  load(file = file.path(dir_data, "extract", nom_ref_bin))
  
  ngrids <- dim(array_obs)[1] # nomber of grids
  scoreprob <- array (NA, dim=c(ngrids,10)) # array for scores to be stored in
  rapport_climps1 <- vector()
  rapport_climps2 <- vector()
  
  # time aggregation
  if (temporalite=="single"){
    seas_obs <- array_obs
    seas_prev <- array_prev
    seas_clim <- array_clim
  } else if (temporalite=="multi") {
    seas_obs <- array()
    seas_prev <- array()
    seas_clim <- array()
    if (agreg=="sum"){
    seas_prev <- apply(array_prev[,,,temp], c(1,2,3), sum) 
    seas_clim <- apply(array_clim[,,,temp], c(1,2,3), sum) 
    seas_obs <- apply(array_obs[,,temp], c(1,2), sum)
    } else if (agreg=="mean"){
    seas_prev <- apply(array_prev[,,,temp], c(1,2,3), mean) 
    seas_clim <- apply(array_clim[,,,temp], c(1,2,3), mean) 
    seas_obs <- apply(array_obs[,,temp], c(1,2), mean)
    }
  }
  
  # Area Under Curve
  scoreprob[,1] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat1 
  scoreprob[,2] <- veriApply("EnsRoca", fcst=seas_clim, obs=seas_obs, prob=proba, parallel = TRUE)$cat1 
  
  # regardin selected quantile, index of the upper one change (5 for quintile, 3 for tercile) 
  if (kantile=="quint"){
    scoreprob[,3] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat5
    scoreprob[,4] <- veriApply("EnsRoca", fcst=seas_clim, obs=seas_obs, prob=proba, parallel = TRUE)$cat5
  } else { 
    scoreprob[,3] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat3 
    scoreprob[,4] <- veriApply("EnsRoca", fcst=seas_clim, obs=seas_obs, prob=proba, parallel = TRUE)$cat3 
    }

  # loop needed for Brierscore function, as veriapply doesn't work on parallelized loop
  for (i in 1:ngrids){
    # Brier score
  scoreprob[i,5] <- BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_ps[i,,1])))$bs
  scoreprob[i,6] <- BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_clim[i,,1])))$bs
    # intermediate for Brier skill score
  rapport_climps1[i] <- BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_ps[i,,1])))$bs/BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_clim[i,,1])))$bs
  
    # Brier score
  scoreprob[i,7] <- BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_ps[i,,events])))$bs
  scoreprob[i,8] <- BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_clim[i,,events])))$bs
    # intermediate for Brier skill score
  rapport_climps2[i] <- BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_ps[i,,events])))$bs/BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_clim[i,,events])))$bs
  
  }
  # Brier skill score
  scoreprob[,9] <- 1-rapport_climps1
  scoreprob[,10] <- 1-rapport_climps2
  # naming file
  nom_scoreroc <- paste0(indicateur,"_", b, scalename(selected_project,temporalite,agreg), kantile, "aucgrid.RData")
  # naming file
  save(scoreprob, file = file.path(dir_scores, nom_scoreroc))

  }
  
stopCluster(cl) # stop parallelizing

}

tictoc::toc() # stop timer


```

## Box-agregated scores

We compute AUC and Brier score at box level, using manual functions defined in "functions"
It requires probabilistic data computed in "transfo_probin"

```{r}

tictoc::tic() # timer on

for (indicateur in indicateurs){

  d <- which(indicateurs==indicateur)
  temporalite <- time_scale[d] # select corresponding time dimension: single or multi
  agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean
  
  auc_boot_ps <- list()
  auc_boot_clim <- list()
  brier_boot_ps <- list()
  brier_boot_clim <- list()
  
  load(file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))
  ngrids <- length(lat)

  for (b in 1:nboot){ 
    
    # loading probabilistic data
  nom_prev_prob <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"psprob",kantile, ".RData")
  nom_clim_prob <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"climprob",kantile, ".RData")
  nom_ref_bin <- paste0(indicateur,"_",b,scalename(selected_project,temporalite,agreg),"refbin",kantile, ".RData")
 
    load(file = file.path(dir_data, "extract", nom_prev_prob))
    load(file = file.path(dir_data, "extract", nom_clim_prob))
    load(file = file.path(dir_data, "extract", nom_ref_bin))
    
    int_auc_ps <- vector()
    int_auc_clim <- vector()
    int_brier_ps <- vector()
    int_brier_clim <- vector()
    for (e in c(1:events)){ # events are different quantiles
      
      int_auc_ps[e] <- score_boite_prob(prob_ps, bin_ref, lat, nb_bin = 10, e)$AUC # auc 
      int_auc_clim[e] <- score_boite_prob(prob_clim, bin_ref, lat, nb_bin = 10, e)$AUC # auc 
      int_brier_ps[e] <- brier_score_boite(prob_ps, bin_ref, lat, e) # brier  
      int_brier_clim[e] <- brier_score_boite(prob_clim, bin_ref, lat, e) # brier  
  
      }
   # we save them in lists to keep all bootstrapped score
   auc_boot_ps[[b]] <- int_auc_ps
   auc_boot_clim[[b]] <- int_auc_clim
   brier_boot_ps[[b]] <- int_brier_ps 
   brier_boot_clim[[b]] <- int_brier_clim 
  
  }
  
  #naming and saving auc list 
  auc_nom_ps <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "boxauc_ps.RData")
  auc_nom_clim <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "boxauc_clim.RData")

  save(auc_boot_ps, file = file.path(dir_scores, auc_nom_ps))
  save(auc_boot_clim, file = file.path(dir_scores, auc_nom_clim))

  #naming and saving brier list 
  brier_nom_ps <- paste0(indicateur,scalename(selected_project,temporalite,agreg), "_", kantile, "boxbrier_ps.RData")
  brier_nom_clim <- paste0(indicateur,scalename(selected_project,temporalite,agreg), "_", kantile, "boxbrier_clim.RData")
  save(brier_boot_ps, file = file.path(dir_scores, brier_nom_ps))
  save(brier_boot_clim, file = file.path(dir_scores, brier_nom_clim))
  
}

tictoc::toc() # timer off

```


## Box-agregated score analysis

From lists generated before, we create and save a csv with bootstrapped score mean (for box score only)

```{r}
df_auc_ps <- data.frame()
df_auc_clim <- data.frame()
df_brier_ps <- data.frame()
df_brier_clim <- data.frame()

for (indicateur in indicateurs){
  
  d <- which(indicateurs==indicateur)
  temporalite <- time_scale[d] # select corresponding time dimension: single or multi
  agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean

  # loading list files
  auc_nom_ps <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "boxauc_ps.RData")
  load(file = file.path(dir_scores, auc_nom_ps))
  auc_nom_clim <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "boxauc_clim.RData")
  load(file = file.path(dir_scores, auc_nom_clim))
  
  brier_nom_ps <- paste0(indicateur,scalename(selected_project,temporalite,agreg), "_", kantile, "boxbrier_ps.RData")
  load(file = file.path(dir_scores, brier_nom_ps))
  brier_nom_clim <- paste0(indicateur,scalename(selected_project,temporalite,agreg), "_", kantile, "boxbrier_clim.RData")
  load(file = file.path(dir_scores, brier_nom_clim))
  
  # averaging list
  df_auc_ps <- array(unlist(auc_boot_ps) , c(events,nboot))
  df_auc_clim <- array(unlist(auc_boot_clim) , c(events,nboot))
  df_brier_ps <- array(unlist(brier_boot_ps) , c(events,nboot))
  df_brier_clim <- array(unlist(brier_boot_clim) , c(events,nboot))

  sum_aucbox_ps <- apply(df_auc_ps, 1, mean, na.rm = T)
  sum_aucbox_clim <- apply(df_auc_clim, 1, mean, na.rm = T)
  sum_brierbox_ps <- apply(df_brier_ps, 1, mean, na.rm = T)
  sum_brierbox_clim <- apply(df_brier_clim, 1, mean, na.rm = T)
  
  # statistical tests
   sum_aucbox_ps[4:6] <- c(wilcox.test(df_auc_ps[1,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value, NA, wilcox.test(df_auc_ps[events,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value) # first and last quantile
   sum_aucbox_clim[4:6] <- c(wilcox.test(df_auc_clim[1,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value, NA, wilcox.test(df_auc_clim[events,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value) # first and last quantile
   sum_brierbox_ps[4:6] <- c(wilcox.test(df_brier_ps[1,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value, NA, wilcox.test(df_brier_ps[events,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value) # first and last quantile
   sum_brierbox_clim[4:6] <- c(wilcox.test(df_brier_clim[1,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value, NA, wilcox.test(df_brier_clim[events,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value) # first and last quantile
  

  # naming and saving files
  nom_aucav_ps <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "avboxauc_ps.csv")
  write_csv(as.data.frame(sum_aucbox_ps), file.path(dir_scores, nom_aucav_ps))
  nom_aucav_clim <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "avboxauc_clim.csv")
  write_csv(as.data.frame(sum_aucbox_clim), file.path(dir_scores, nom_aucav_clim))

  nom_brierav_ps <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "avboxbrier_ps.csv")
  write_csv(as.data.frame(sum_brierbox_ps), file.path(dir_scores, nom_brierav_ps))
  nom_brierav_clim <- paste0(indicateur,"_", scalename(selected_project,temporalite,agreg), kantile, "avboxbrier_clim.csv")
  write_csv(as.data.frame(sum_brierbox_clim), file.path(dir_scores, nom_brierav_clim))
  
}

```

