---
title: "Deterministic scores"
output: html_document
---

```{r setup, include=FALSE}

dir_data <- "D:/HOME/ekamir/scoring_medscope/data/stics" 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure"

library(dplyr)
library(tidyverse)
library(plyr)
library(easyVerification)
library(verification)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)
library(readr)

```
## Parameters 

```{r }

# temporal agregation
temp <- 9:20

# bootstrap and cores
no_core <- 6
nboot <- 100

# variables
indicateurs <- c("drainall", "etall", "yieldwheat", "yieldgrass", "yieldgrassC2C3")
  ### temporalité
  if (indicateur=="yieldwheat"|indicateur=="yieldgrass"|indicateur=="yieldgrassC2C3"){
      temporalite <- "yearly"
  }else {temporalite <- "quinzaine"}

```

                                            ### GRIDDED ###
                                            
                                            
## Gridded scores

Scores computed for each grid separately.
Indicators are assessed on a 28 days time window and over the whole season

```{r}

tictoc::tic()

acc <- vector()
rmse_ps <- vector()
rmse_clim <- vector()

for (indicateur in indicateurs){

  cl <- makeCluster(no_core)
  registerDoParallel(cl)

  foreach(b=1:nboot, .packages=c('easyVerification', 's2dverification')) %dopar%  { # loop on bootstrap samples
  #Spearman <- easyVerification:::EnsCorr
    
    nom_ref <- paste0(indicateur,b,"_ref_uerra5.RData")
    nom_prev <- paste0(indicateur,b,"_prev_uerra5.RData")
    nom_clim <- paste0(indicateur,b,"_clim_uerra5.RData")
    nom_an_prev <- paste0(indicateur,b,"_anomaly_prev_uerra5.RData")
    nom_an_obs <- paste0(indicateur,b,"_anomaly_obs_uerra5.RData")
    
    load(file = file.path(dir_data, "extract", nom_ref))
    load(file = file.path(dir_data, "extract", nom_prev))
    load(file = file.path(dir_data, "extract", nom_clim))
    load(file = file.path(dir_data, "extract", nom_an_prev))
    load(file = file.path(dir_data, "extract", nom_an_obs))
    
    ngrids <- dim(array_obs)[1]
    det_seas <- array(NA, dim = c(ngrids,3))
    
  if (temporalite=="yearly"){
    prev_seas <- array_prev 
    clim_seas <- array_clim
    ref_seas <- array_obs
  } else if (temporalite=="quinzaine") {
          prev_seas <- apply(array_prev[,,,temp], c(1,2,3), sum) 
          clim_seas <- apply(array_clim[,,,temp], c(1,2,3), sum)
          ref_seas <- apply(array_obs[,,temp], c(1,2), sum)
    }
    
    
      # RMSE #
 
    rmse_ps <- veriApply("EnsRmse", fcst=prev_seas, obs=ref_seas, parallel = TRUE)
    rmse_clim <- veriApply("EnsRmse", fcst=clim_seas, obs=ref_seas, parallel = TRUE)
          
      # ACC #
    
    acc <- veriApply("EnsCorr", fcst=prev_anomaly, obs=obs_anomaly, parallel = TRUE)
    
    score_det <- array(c(acc, rmse_ps, rmse_clim), dim=c(ngrids,3))
    
    nom_det <- paste0(indicateur, b,"scoredet_grid_uerra5.RData")
  
    save(score_det, file = file.path(dir_data, "score", nom_det))
  

}
stopCluster(cl)

}
tictoc::toc() 



```

  
                                            ### BOX SCORES ###
                                            
                                            
## Aggregated scores

Scores computed for each grid separately.
Indicators are assessed on a 28 days time window and over the whole season

```{r}

coeff_rmsps <- vector()
coeff_rmsclim <- vector()
coeff_acc <- vector()

for (indicateur in indicateurs){
  
  load(file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))

  det_boot <- data.frame()

  for(b in 1:nboot)   {
    
    # loading gridded scores
  
    nom_prevseas <- paste0(indicateur, b,"scoredet_grid_uerra5.RData")
    load(file = file.path(dir_data, "score", nom_prevseas))
    ngrids <- dim(score_det)[1]

    val_prevseas <- list()
  
    for (i in 1:ngrids){
      val_prevseas[[i]] <- score_det[i,]*cos(pi*lat[i]/180)
      coeff_rmsps[i] <- cos(pi*lat[i]/180)
      coeff_rmsclim[i] <- cos(pi*lat[i]/180)
      coeff_acc[i] <- cos(pi*lat[i]/180)
      
     # removing NAs
      if (is.na(score_det[i,1])==TRUE){
        coeff_acc[i] <- NA
      }
      
      if (is.na(score_det[i,2])==TRUE){
        coeff_rmsps[i] <- NA
      }
      
      if (is.na(score_det[i,3])==TRUE){
        coeff_rmsclim[i] <- NA
      }

          }
    
    # grid average
    array_score <- apply(array(unlist(val_prevseas) , c(3,ngrids)), 1, sum, na.rm = T)
    det_boot[b,1] <- array_score[1]/sum(coeff_acc, na.rm = T)
    det_boot[b,2] <- array_score[2]/sum(coeff_rmsps, na.rm = T)
    det_boot[b,3] <- array_score[3]/sum(coeff_rmsclim, na.rm = T)
  }
  
    nom_bootseas <- paste0(indicateur, "scoredet_boot_uerra5.csv")
    save(det_boot, file = file.path(dir_data, "score", nom_bootseas))
    
    det_meanboot <- apply(det_boot, 2, mean, na.rm=T)
    #colnames(det_meanboot) <- c("msess","acc")
    nom_meanseas <- paste0(indicateur, "detbox_meanboot_uerra5.csv")
    write_csv(as.data.frame(det_meanboot), file.path(dir_data, "score", nom_meanseas))

}

```


## Box score significance test

Wilcoxon test with H0: MSESS non supérieure à 0 + ACC non supérieur à 0.3

```{r}
ind <- 2

   # season
  nom_bootseas <- paste0(indicateur, "scoredet_boot.csv")
  load(file = file.path(dir_data, "score", nom_bootseas))

  boxboot <- array(unlist(det_boot) , c(2,nboot))
  
  wilcox.test(boxboot[1,], mu = 0, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value # msess
  wilcox.test(boxboot[2,], mu = 0.3, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value # acc


```
