---
title: "Area under ROC"
output: html_document
---

```{r setup, include=FALSE}

dir_data <- "D:/HOME/ekamir/scoring_medscope/data/stics" 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure" 

library(easyVerification)
library(verification)
library(dplyr)
library(stats)
library(SpecsVerification)
library(s2dverification)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)
library(readr)
library(purrr)

```

## Parameters 

```{r }

# temporal agregation
temp <- 9:20

# chosen quantile
kantile <- "ter" # ter or quint

# bootstrap and cores
no_core <- 6
nboot <- 100

if (kantile=="quint"){
  proba = 1:4/5; events = 5
} else { proba = 1:2/3; events = 3}

# variables
indicateurs <- c("drainall", "etall", "yieldwheat", "yieldgrass", "yieldgrassC2C3")


```
                                            
## Gridded scores

Scores computed for each grid separately.
Indicators are assessed on a 28 days time window and over the whole season

1: roc quant1
2: roc last quant
3: brier quant1
4: brier ss quant1
5: brier last quant
6: brier ss last quant

```{r}

tictoc::tic()

for (indicateur in indicateurs){
  
  # temporalité
  if (indicateur=="yieldwheat"|indicateur=="yieldgrass"|indicateur=="yieldgrassC2C3"){
      temporalite <- "yearly"
  }else {temporalite <- "quinzaine"}

  cl <- makeCluster(no_core)
  registerDoParallel(cl)

foreach(b=1:nboot, .packages = c('easyVerification', 's2dverification'), .combine=rbind) %dopar%  { # loop on bootstrap samples


  nom_ref <- paste0(indicateur,b,"_ref_uerra5.RData")
  nom_prev <- paste0(indicateur,b,"_prev_uerra5.RData")
  nom_prev_prob <- paste0(indicateur,b,"psprob_uerra5",kantile, ".RData")    
  nom_clim_prob <- paste0(indicateur,b,"climprob_uerra5",kantile, ".RData")    
  nom_ref_bin <- paste0(indicateur,b,"refbin_uerra5",kantile, ".RData")
  
  load(file = file.path(dir_data, "extract", nom_prev))
  load(file = file.path(dir_data, "extract", nom_ref))
  load(file = file.path(dir_data, "extract", nom_prev_prob))
  load(file = file.path(dir_data, "extract", nom_clim_prob))
  load(file = file.path(dir_data, "extract", nom_ref_bin))
  
  
  ngrids <- dim(array_obs)[1]
  scoreprob <- array (NA, dim=c(ngrids,6))
  rapport_climps1 <- vector()
  rapport_climps2 <- vector()
  
  
  if (temporalite=="yearly"){
    seas_obs <- array_obs
    seas_prev <- array_prev
  } else if (temporalite=="quinzaine") {
    seas_obs <- array()
    seas_prev <- array()
    seas_obs <- apply(array_obs[,,temp],c(1,2), sum)# seasonal: WARNING WITH AGREGATION METHOD!!!!
    seas_prev <- apply(array_prev[,,,temp],c(1,2,3), sum)
    
  }
  
  scoreprob[,1] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat1 
  
  if (kantile=="quint"){
    scoreprob[,2] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat5
  } else { scoreprob[,2] <- veriApply("EnsRoca", fcst=seas_prev, obs=seas_obs, prob=proba, parallel = TRUE)$cat3 }

  for (i in 1:ngrids){
  scoreprob[i,3] <- BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_ps[i,,1])))$bs
  rapport_climps1[i] <- BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_ps[i,,1])))$bs/BrierScore(as.matrix(t(bin_ref[i,,1])), as.matrix(t(prob_clim[i,,1])))$bs
  

  scoreprob[i,5] <- BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_ps[i,,events])))$bs
  rapport_climps2[i] <- BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_ps[i,,events])))$bs/BrierScore(as.matrix(t(bin_ref[i,,events])), as.matrix(t(prob_clim[i,,events])))$bs
  
  }
  scoreprob[,4] <- 1-rapport_climps1
  scoreprob[,6] <- 1-rapport_climps2
  nom_scoreroc <- paste0(indicateur, b, "_", kantile, "aucgrid_uerra5.RData")

  save(scoreprob, file = file.path(dir_data, "score", nom_scoreroc))

  }
  
stopCluster(cl)

}

tictoc::toc()


```

## Box-agregated scores

Aggregation computed using latitude weighting   

```{r}

tictoc::tic()

for (indicateur in indicateurs){

  auc_boot <- list()
  brier_boot <- list()
  
  load(file = file.path(dir_data, "extract", paste0(indicateur, "_lat.RData")))
  ngrids <- length(lat)

  for (b in 1:nboot){
    
    nom_prev <- paste0(indicateur,b,"psprob_uerra5",kantile, ".RData")    
    nom_ref <- paste0(indicateur,b,"refbin_uerra5",kantile, ".RData")
    load(file = file.path(dir_data, "extract", nom_prev))
    load(file = file.path(dir_data, "extract", nom_ref))
    
    int_auc <- vector()
    int_brier <- vector()
    for (e in c(1:events)){ # through different quantiles
      
      int_auc[e] <- score_boite_prob(prob_ps, bin_ref, lat, nb_bin = 10, e)$AUC # seasonal 
      int_brier[e] <- brier_score_boite(prob_ps, bin_ref, lat, e) # seasonal 
  
      }
 
   auc_boot[[b]] <- int_auc 
   brier_boot[[b]] <- int_brier 
  
  }
 
  nom_box <- paste0(indicateur, "_", kantile, "boxauc_uerra5.RData")
  save(auc_boot, file = file.path(dir_data, "score", nom_box))

  brier_box <- paste0(indicateur, "_", kantile, "boxbrier_uerra5.RData")
  save(brier_boot, file = file.path(dir_data, "score", brier_box))
  
}

tictoc::toc()

```


## Box-agregated score analysis

Transform lists in a more handy format

```{r}

for (indicateur in indicateurs){
  
   # season
  nom_auc <- paste0(indicateur, "_", kantile, "boxauc_uerra5.RData")
  load(file = file.path(dir_data, "score", nom_auc))
  nom_brier <- paste0(indicateur, "_", kantile, "boxbrier_uerra5.RData")
  load(file = file.path(dir_data, "score", nom_brier))
  
  sum_aucbox <- apply(array(unlist(auc_boot) , c(events,nboot)), 1, mean, na.rm = T)
  sum_brierbox <- apply(array(unlist(brier_boot) , c(events,nboot)), 1, mean, na.rm = T)
  
  nom_aucav <- paste0(indicateur, "_", kantile, "avboxauc_uerra5.csv")
  write_csv(as.data.frame(sum_aucbox), file.path(dir_data, "score", nom_aucav))
  nom_brierav <- paste0(indicateur, "_", kantile, "avboxbrier_uerra5.csv")
  write_csv(as.data.frame(sum_aucbox), file.path(dir_data, "score", nom_brierav))
  
}

```

## Box score significance test

Wilcoxon test with H0: AUC non supérieure à 0.5

```{r}
indicateur <- "bilanhydr"
indicateur <- "grassyieldC1C2_"
lastquant <- 3 # 3 if terciles or 5

  nom_box <- paste0(indicateur, "_", kantile, "boxauc_uerra5.RData")
  load(file = file.path(dir_data, "score", nom_box))
  
  boxboot <- array(unlist(auc_boot) , c(events,nboot))
  
   wilcox.test(boxboot[1,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value
   wilcox.test(boxboot[lastquant,], mu = 0.5, alternative = "greater", conf.level = 0.95, na.action = na.omit)$p.value


```
