---
title: "Probabilistic transformation"
output: html_document
---

This code aims at computing respectively probabilistic and binary data from "raw" data (extracted by "extraction") of prevision/climato and reference. 

```{r setup, include=FALSE}

library(dplyr)
library(stats)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)
library(easyVerification)

```

## Parameters 

Some are loaded (because they remain permanent through codes)
Some are set here and can therefore be modified by user

```{r}

                                     ### SELECTED PARAMETERS (to be set by user) ###

# selected project and directories
selected_project <- "stics" # stics or season or tree
dir_data <- paste0("D:/HOME/ekamir/scoring_medscope/data/",selected_project) 
dir_plot <- "D:/HOME/ekamir/scoring_medscope/figure" 

# number of cores
no_core <- 6

# chosen quantile
kantile <- "ter" # ter or quint
if (kantile=="quint"){
  proba = 1:4/5; events = 5
} else { proba = 1:2/3; events = 3}


                                      ### LOADED PARAMETERS (permanent) ###


# loading number of bootstrap interations
load(file = file.path(dir_data, "extract", "nboot.RData"))

# loading variables
load(file = file.path(dir_data, "extract", "indicateurs.RData"))
# loading time scales and agregation type
load(file = file.path(dir_data, "extract", "temporalite.RData")) # is there a time dimension
load(file = file.path(dir_data, "extract", "time_window.RData")) # if yes which window is selected
load(file = file.path(dir_data, "extract", "agregation.RData")) # if yes what type of aggregation should be applied

ntemp <- length(temp) # number of elements in time dimension

# loading ensemble members
load(file = file.path(dir_data, "extract", "runs.RData"))
nrun <- length(run) # number of members

# loading years
load(file = file.path(dir_data, "extract", "years.RData"))
nan <- length(an) # number of years

```

## Transforming data 

Regarding selected event (tercile/quintile) reference continuous values are transformed into binary values; ensemble prevision and climatology values are transformed into probabilistic values (computed from ensemble members)

```{r}

tictoc::tic() # timer on

load(file = file.path(dir_data, "extract", "years.RData")) # list of years in each bootstrap iteration

for (indicateur in indicateurs){
  
  d <- which(indicateurs==indicateur)
  temporalite <- time_scale[d] # select corresponding time dimension: single or multi
  agreg <- agreg_type[d] # select corresponding time aggregation type: sum or mean

cl <- makeCluster(no_core) # setting up number of cores
registerDoParallel(cl) # parallelizing on them

foreach(b=1:nboot, .packages = 'easyVerification') %dopar%  { # for-loop parallelized
  
  # loading "raw" data
  nom_ref <- paste0(indicateur,b,"_ref_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_ref))
  nom_prev <- paste0(indicateur,b,"_prev_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_prev))
  nom_clim <- paste0(indicateur,b,"_clim_uerra5.RData")
  load(file = file.path(dir_data, "extract", nom_clim))

  ngrids <- dim(array_obs)[1]

  # time aggregation
  if (temporalite=="single"){
    int_obs <- array_obs 
    int_prev <- array_prev
    int_clim <- array_clim
  } else if (temporalite=="multi") {
    if (agreg=="sum"){
    int_obs <- apply(array_obs[,,temp],c(1,2), sum) 
    int_prev <- apply(array_prev[,,,temp],c(1,2,3), sum)
    int_clim <- apply(array_clim[,,,temp],c(1,2,3), sum)
    } else if (agreg=="mean"){
    int_obs <- apply(array_obs[,,temp],c(1,2), mean) 
    int_prev <- apply(array_prev[,,,temp],c(1,2,3), mean)
    int_clim <- apply(array_clim[,,,temp],c(1,2,3), mean)
    }
  }
  
  # setting up arrays for data to be stored in
  bin_ref <- array (NA, dim=c(ngrids,nan,events))
  prob_ps <- array (NA, dim=c(ngrids,nan,events))
  prob_clim <- array (NA, dim=c(ngrids,nan,events))
  
  # using for-loop on grids, probabilistic values are derived using two nested functions from EasyVerification
  for (i in 1:ngrids){
    bin_ref[i,,] <- count2prob(convert2prob(int_obs[i,], prob = proba), type = 4)
    prob_ps[i,,] <- count2prob(convert2prob(int_prev[i,,], prob = proba), type = 4)
    prob_clim[i,,] <- count2prob(convert2prob(int_clim[i,,], prob = proba), type = 4)
  }
  
  # naming and saving files
  nom_prev <- paste0(indicateur,b,"psprob_uerra5",kantile, ".RData")
  nom_clim <- paste0(indicateur,b,"climprob_uerra5",kantile, ".RData")
  nom_ref <- paste0(indicateur,b,"refbin_uerra5",kantile, ".RData")
    
  save(prob_ps, file = file.path(dir_data, "extract", nom_prev))
  save(prob_clim, file = file.path(dir_data, "extract", nom_clim))
  save(bin_ref, file = file.path(dir_data, "extract", nom_ref))
} 
 
  }

tictoc::toc() # timer off

```
